Using \LaTeX\ to submit manuscripts to peer reviewed journals.

### 2014-02-11 docs/thesis/chap-theoretical_background.tex
Wrote okapi/bm25 section

>>>>

\section{Probabilistic Relevance Model}
The probabilistic relevance model was developed in 1976 by Robertson and Jones\cite{probabilistic_relevance_model}. The similarity of a document $d_{j}$ to a query $q$ is given by the probability of $d_{j}$ being relevant to $q$. It assumes there is a set $R$ that is preferred to be the answer set for $q$. The model assumes that the document in $R$ are relevant to $q$ and the documents in $\bar{R}$ are not relevant. $R$ then is the set that maximizes:

$$sim(d_{j},q) = \frac{ P(R|d_{j})} {P(\bar{R} |d_{j})} $$

The importance of this model is that it serves as a solid base for state of the art information retrieval models.

\section{Okapi BM25}

The Okapi BM25 ranking function is one of the current state-of-the-art models in information retrieval\cite{okapibm25}. The actual scoring function is called BM25 and Okapi refers to the first system implementing this function in the 1980s. 
Given a query $Q$ as a vector of keywords $q_{1}, q_{2}, ... , q_{n}$ then the score for a document $d$ is:

$$score(d, Q) = \sum_{i=1}^{n}IDF{(q_i)}*\frac{ frec(q_i, D)}{frec(q_i, D)+k_1*((1-b)+b*\frac{|D|}{avgdl})} $$
and $$IDF(q_i) = \log{\frac{N - df(q_i)+0.5}{df(q_i)+0.5}} $$

where $frec(q_i, D)$ is the number of occurrences of keyword $q_i$ in $D$, $avgdl$ is the average length (In keywords) of a document, and $k_1$ and $b$ are free parameters. Usually $k_1 = 2$ and $b \in [0, 1]$ controls how important is the length normalization, being set normally to $0.75$. $N$ is the size of the collection, $df(q_i)$ is the number of documents that contain term $q_i$

A further variation that takes into account structure in the form of a document containing different fields and each of them having its own length and importance (boost) can be formulated as follows, known as BM25F can be expressed as follows:

$$score(d, Q) = \sum_{i=1}^{n} \frac{weigth(q_i, d)}{k+ weigth(q_i, d)} * IDF(q_i)$$
and 
$$weight(q_i, d) = \sum_{j=1}^m \frac{frec(q_i, D, field_j)*boost_j)}{((1-b_j) + b_j*\frac{|l_j|}{avgfl_j})} $$
where $m$ is the number of different fields, $boost_j$ is the relative importance for each field, $b_j$ is analogous to the $b$ constant in the BM25 group of equations, $|l_j|$ is the length of the $field_j$ and $avgfl_j$ is the average length of the $field_j$.
